---
title: "VMH LepRb-Vglut2 hM3Dq CLAMS"
author: "Alan Rupp"
date: "`r strftime(Sys.time(), '%b %d, %Y')`"
output: pdf_document
geometry: margin=0.5in
---

Plotting and analyzing CLAMS data from the Core for Vglut2-Flp;Lepr-Cre mice that Paul injected with his Flp+Cre hM3Dq virus.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(bitmapType = "cairo")
library(tidyverse)
library(readxl)
library(lme4); library(emmeans)
library(knitr); library(kableExtra)
library(lubridate)
```


```{r read}
fname <- "2019-09-15 Myers-Paul 9 mice 5 day run CLAMS mt.xlsm"
sheets <- excel_sheets(fname)

read_data <- function(sheet_name) {
  df <- read_xlsx(fname, sheet = sheet_name, skip = 59)
  df <- filter(df, !Computer %in% c("Date/Clock", "Time"))
  df <- select(df, -starts_with(".."), -`Hourly Average`)
  df <- select(df, -`797`) # 797 is an injection miss
  df <- rename(df, "Date" = Computer, "Time" = `ID#:`)
  df <- mutate(df, Date = as.Date(as.numeric(Date), 
                                  origin = "1900-01-01") - 2)
  df <- mutate(df, Time = as.numeric(Time))
  df <- mutate(df, Time = ifelse(Time >= 1, Time - as.integer(Time), Time))
  df <- gather(df, -Date, -Time, key = "Mouse", value = "value")
  df <- mutate(df, "treatment" = ifelse(Date <= as.Date("2019-09-17"),
                                        "Saline", "CNO"))
  df <- mutate(df, treatment = factor(treatment, levels = c("Saline", "CNO")))
  df <- mutate(df, "tod" = ifelse(Time > 0.25 & Time <= 0.75, "Day", "Night"))
}

to_read <- c("VO2" = "VO2-mouse ", "VCO2" = "VCO2-mouse", "RER" = "RER",
             "EE" = "EE-mouse", "Activity-X" = "Activity-X total",
             "Activity-Z" = "Activity-Z total",
             "Ambulatory" = "Activity-X Ambulatory",
             "Fat Oxidation" = "Fat Oxidation-LBM",
             "Glucose Oxidation" = "Glucose Oxidation-LBM",
             "Food Intake" = "Food Intake")

df <- map(to_read, read_data)
names(df) <- names(to_read)
```

```{r summarize-data}
summarize_data <- function(parameter) {
  df <- df[[parameter]] %>%
    filter(Date != as.Date("2019-09-15") & Date != as.Date("2019-09-20")) %>%
    group_by(Time, Mouse, treatment) %>%
    summarize(value = mean(value, na.rm = TRUE)) %>%
    group_by(Time, treatment) %>%
    summarize(avg = mean(value, na.rm = TRUE), 
              sem = sd(value, na.rm = TRUE) / sqrt(n()))
  return(df)
}
```

## Overview
There are `r length(unique(df[["EE"]]$Mouse))` mice that were subjected to the regimen of 2 days of 2 saline injections, followed by 2 days of 2 CNO injections. 

```{r}
df[["EE"]] %>% group_by(Mouse, treatment, Date) %>% 
  count() %>%
  filter(n > 70) %>%
  select(-n) %>%
  spread(key = "Date", value = "treatment") %>%
  kable(caption = "Experiment full days") %>%
  kable_styling(latex_options = "hold_position")
```

Injections were performed at 9:30am and 4:30pm.

\pagebreak

## Averaged data
The full data, averaged across both days of treatment showing the mean and standard error. Dotted lines correspond to the injection times and the shaded gray corresponds to the light cycle.

```{r plot-continuous, fig.width = 7.5, fig.height = 9}
plot_continuous <- function(parameter, ylabel = NULL, no_title = FALSE,
                            legend = "right") {
  # get mean and error
  df <- summarize_data(parameter)
  # plot settings
  colors <- c("gray", "red3")
  legend_title <- expression(underline("Treatment"))
  # plot
  p <- ggplot(df, aes(x = Time, y = avg, color = treatment)) +
    annotate("rect", xmin = 0, xmax = 0.25, 
             ymin = min(df$avg - df$sem), ymax = max(df$avg + df$sem), 
             alpha = 0.3, fill = "gray") +
    annotate("rect", xmin = 0.75, xmax = 1, 
             ymin = min(df$avg - df$sem), ymax = max(df$avg + df$sem), 
             alpha = 0.3, fill = "gray") +
    geom_line() +
    geom_ribbon(aes(ymin = avg - sem, ymax = avg + sem, fill = treatment),
                alpha = 0.3, color = 0) +
    scale_color_manual(values = colors, name = legend_title) +
    scale_fill_manual(values = colors, name = legend_title) +
    geom_vline(aes(xintercept = 9.5/24), linetype = "dashed") +
    geom_vline(aes(xintercept = 16.5/24), linetype = "dashed") +
    theme_classic() +
    ggtitle(parameter) +
    ylab(NULL) + xlab("Time of day") +
    scale_x_continuous(expand = c(0, 0), limits = c(0, 1),
                       breaks = seq(0, 1, by = 0.25),
                       labels = c("00:00", "6:00", "12:00", "18:00","24:00")) +
      scale_y_continuous(expand = c(0, 0)) +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text = element_text(color = "black"),
          legend.position = legend)
  if (!is.null(ylabel)) p <- p + ylab(ylabel) 
  if (no_title) p <- p + theme(plot.title = element_blank())
  return(p)
}

cowplot::plot_grid(plotlist = map(names(df), plot_continuous), ncol = 2)
```

\pagebreak

## Paired summary
This is the data averaged by treatment with pairing because each mouse was given each treatment (Saline and CNO).

```{r plot-paired, fig.width = 7.5, fig.height = 9}
plot_paired <- function(parameter, daily = FALSE,
                        no_title = FALSE, no_ticks = FALSE,
                        no_text = FALSE, ylabel = NULL) {
  df <- df[[parameter]] %>%
    filter(Date != as.Date("2019-09-15") & Date != as.Date("2019-09-20"))
  if (daily) {
    df <- df %>% group_by(Mouse, treatment, day(Date)) %>%
      summarize(value = sum(value, na.rm = TRUE)) %>%
      group_by(Mouse, treatment) %>%
      summarize(value = mean(value))
  } else {
    df <- df %>% group_by(Mouse, treatment) %>%
      summarize(value = mean(value, na.rm = TRUE))
  }
  # group averages
  group_df <- df %>%
    group_by(treatment) %>%
    summarize(avg = mean(value, na.rm = TRUE),
              sem = sd(value, na.rm = TRUE) / sqrt(n()))
  # plot
  p <- ggplot(df, aes(x = treatment, y = value)) +
    geom_col(data = group_df, aes(y = avg, fill = treatment), 
             color = "black", position = "dodge") +
    geom_errorbar(data = group_df, 
                  aes(x = treatment, ymin = avg - sem, ymax = avg + sem), 
                  width = 0.4, inherit.aes = FALSE, 
                  position = position_dodge(width = 0.9)) +
    geom_line(aes(group = Mouse)) +
    geom_point() +
    scale_fill_manual(values = c("gray", "red3"),
                      name = expression(underline("Treatment"))) +
    theme_classic() +
    ylab(NULL) + xlab(NULL) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(parameter) +
    theme(legend.position = "none",
          axis.text = element_text(color = "black"),
          axis.text.x = element_text(size = 10),
          strip.background = element_blank(),
          plot.title = element_text(hjust = 0.5),
          legend.background = element_blank())
  # get stats (paired t-test)
  result <- t.test(df[df$treatment == "CNO", ]$value,
                   df[df$treatment == "Saline", ]$value,
                   paired = TRUE)$p.value
  # set ymin and ymax values
  ymax <- max(df$value, na.rm = TRUE)*1.05
  if (parameter %in% c("VO2", "VCO2", "RER", "EE")) {
    ymin <- min(df$value, na.rm = TRUE) * 0.95
  } else {
    ymin <- 0
  }
  # add stars to plot
  if (result < 0.05) {
    # different y values based on whether there is a y-axis that goes to 0
    results <- data.frame(x1 = 1, x2 = 2, xstr = 1.5)
    if (parameter %in% c("VO2", "VCO2", "RER", "EE")) {
      results <- results %>% mutate(
        "ybar" = (max(df$value)*1.05-min(df$value)*0.95)*0.05 + max(df$value),
        "ystr" = (max(df$value)*1.05-min(df$value)*0.95)*0.08 + max(df$value)
      )
    } else {
      results <- results %>% mutate(
        "ybar" = max(df$value)*1.05, "ystr" = max(df$value)*1.08
        )
    }
    results$stars <- ifelse(result < 0.001, "***",
                            ifelse(result < 0.01, "**",
                                   ifelse(result < 0.05, "*", NA)))
    # add stars and line to the plot
    p <- p + 
      geom_segment(data = results, 
                   aes(x = x1, xend = x2, y = ybar, yend = ybar)) +
      geom_text(data = results, aes(x = xstr, y = ystr, label = stars))
    # set new ymax to accomodate stars
    ymax <- ifelse(ymax < max(results$ystr)*1.05,
                   max(results$ystr)*1.05,
                   ymax)
  }
  # set limits
  p <- p + coord_cartesian(ylim = c(ymin, ymax))
  # remove elements for simple design
  if (no_title) p <- p + theme(plot.title = element_blank())
  if (no_ticks) p <- p + theme(axis.ticks.x = element_blank())
  if (no_text) p <- p + theme(axis.text.x = element_blank())
  # return plot
  return(p)
}
  
cowplot::plot_grid(plotlist = map(names(df), plot_paired), ncol = 3)
```

\pagebreak

## Paired summary - time of day
Same as previous, but now breaking out by time-of-day (Day v. Night).

```{r plot-paired-tod, fig.width = 7.5, fig.height = 9}
plot_paired_tod <- function(parameter, daily = FALSE, 
                            no_title = FALSE, no_ticks = FALSE,
                            no_text = FALSE) {
  df <- df[[parameter]] %>%
    filter(Date != as.Date("2019-09-15") & Date != as.Date("2019-09-20"))
  if (daily) {
    df <- df %>% group_by(Mouse, treatment, day(Date), tod) %>%
      summarize(value = sum(value, na.rm = TRUE)) %>%
      group_by(Mouse, treatment, tod) %>%
      summarize(value = mean(value))
  } else {
    df <- df %>% group_by(Mouse, treatment, tod) %>%
      summarize(value = mean(value, na.rm = TRUE))
  }
  # average by group for column and errorbar
  group_df <- df %>% group_by(treatment, tod) %>%
    summarize(avg = mean(value, na.rm = TRUE),
              sem = sd(value, na.rm = TRUE) / sqrt(n()))
  # get stats (using a linear mixed model this time)
  model <- lmer(value ~ treatment * tod + (1|Mouse), data = df)
  results <- pairs(emmeans(model, ~ treatment | tod))
  results <- as.data.frame(results)
  # plot
  p <- ggplot(df, aes(x = treatment, y = value)) +
    geom_col(data = group_df, aes(y = avg, fill = treatment), 
             color = "black", position = "dodge") +
    geom_errorbar(data = group_df, 
                  aes(x = treatment, ymin = avg - sem, ymax = avg + sem), 
                  width = 0.4, inherit.aes = FALSE, 
                  position = position_dodge(width = 0.9)) +
    geom_line(aes(group = Mouse)) +
    geom_point() +
    scale_fill_manual(values = c("gray", "red3"),
                      name = expression(underline("Treatment"))) +
    theme_classic() +
    ylab(NULL) + xlab(NULL) +
    scale_y_continuous(expand = c(0, 0)) +
    ggtitle(parameter) +
    theme(legend.position = "none",
          axis.text = element_text(color = "black"),
          strip.background = element_blank(),
          plot.title = element_text(hjust = 0.5),
          legend.background = element_blank(),
          strip.placement = "outside") +
    facet_wrap(~tod, strip.position = "bottom")
  # set ymin and ymax values
  ymax <- max(df$value, na.rm = TRUE)*1.05
  if (parameter %in% c("VO2", "VCO2", "RER", "EE")) {
    ymin <- min(df$value, na.rm = TRUE) * 0.95
  } else {
    ymin <- 0
  }
  # add stars to plot
  if (any(results$p.value < 0.05)) {
    results <- filter(results, p.value < 0.05)
    results <- inner_join(results, df, by = "tod")
    # different y values based on whether there is a y-axis that goes to 0
    if (parameter %in% c("VO2", "VCO2", "RER", "EE")) {
      results <- results %>% group_by(tod, p.value) %>% 
        summarize("ybar" = (max(value)*1.05-min(value)*0.95)*0.1 + max(value),
                  "ystr" = (max(value)*1.05-min(value)*0.95)*0.12 +
                    max(value))
    } else {
      results <- results %>% group_by(tod, p.value) %>% 
        summarize("ybar" = (max(value)*1.05)*0.1 + max(value),
                  "ystr" = (max(value)*1.05)*0.12 + max(value))
    }
    # turn p value into stars
    results <- results %>% mutate(
      "stars" = ifelse(p.value < 0.001, "***",
                               ifelse(p.value < 0.01, "**",
                                      ifelse(p.value < 0.05, "*", NA)))
    )
    # add mapping info to the data.frame
    p <- p + 
      geom_segment(data = results, 
                   aes(x = 1, y = ybar, xend = 2, yend = ybar)) +
      geom_text(data = results, aes(x = 1.5, y = ystr, label = stars),
                inherit.aes = FALSE)
    # set new ymax to accomodate stars
    ymax <- ifelse(ymax < max(results$ystr)*1.05, 
                   max(results$ystr)*1.05, 
                   ymax)
  }
  # set limits
  p <- p + coord_cartesian(ylim = c(ymin, ymax))
  # arguments for minimal design
  if (no_title) p <- p + theme(plot.title = element_blank())
  if (no_ticks) p <- p + theme(axis.ticks.x = element_blank())
  if (no_text) p <- p + theme(axis.text.x = element_blank())
  return(p)
}
  
cowplot::plot_grid(plotlist = map(names(df), plot_paired_tod), ncol = 3)
```

\pagebreak

# Figure
Prepping a figure for the manuscript based on these plots.

```{r}
# this puts the 3 different data plots together
figure_plot <- function(parameter, ylabel = NULL, daily = FALSE) {
  cowplot::plot_grid(plotlist = list(
    plot_continuous(parameter, legend = "none", no_title = TRUE,
                    ylabel = ylabel) +
      theme(axis.text = element_text(size = 8),
            axis.title.x = element_text(size = 10)),
    plot_paired(parameter, daily = daily, no_text = TRUE, no_title = TRUE,
                no_ticks = TRUE) +
      theme(axis.text.y = element_text(size = 8),
            plot.margin = margin(t = 0.1, b = 0.37, unit = "in")),
    plot_paired_tod(parameter, daily = daily, no_text = TRUE, no_title = TRUE,
                    no_ticks = TRUE) +
      theme(axis.text.y = element_text(size = 8))
  ), rel_widths = c(0.5, 0.2, 0.3), ncol = 3)
}
```

```{r}
# this adds a title and line to the data plots
panel_plot <- function(parameter, left_ylabel = NULL, title = NULL,
                       bar_daily = FALSE) {
  if (is.null(title)) title <- parameter
  cowplot::plot_grid(plotlist = list(
    ggplot() +
      annotate("text", x = 1, y = 1, label = title, hjust = 0.5, vjust = 1) +
      annotate("segment", x = 0, xend = 2, y = 0.5, yend = 0.5) +
      theme_void(),
    figure_plot(parameter, ylabel = left_ylabel, daily = bar_daily)
    ), ncol = 1, rel_heights = c(0.1, 0.9))
}
```

```{r, fig.width = 7.5, fig.height = 5.5}
# combining all parameters into a final figure
cowplot::plot_grid(plotlist = list(
  panel_plot("VO2", left_ylabel = "mL/kg/hr", title = expression("VO"[2])),
  panel_plot("VCO2", left_ylabel = "mL/kg/hr", title = expression("VCO"[2])),
  panel_plot("Ambulatory", left_ylabel = "Counts/hr", bar_daily = TRUE),
  panel_plot("EE", left_ylabel = "kcal/hr", title = "Energy Expenditure"),
  panel_plot("Food Intake", left_ylabel = "grams", bar_daily = TRUE),
  panel_plot("RER", left_ylabel = expression("VCO"[2]*"/VO"[2]))
), ncol = 2, labels = letters[6:11], label_fontface = "bold", vjust = 1)
```

\pagebreak

# Files
Saving workspace image as `clams.Rdata`.
```{r}
save.image("clams.Rdata")
```

# Session info
```{r}
sessionInfo()
```